

```{r}
library(sf)
library(ggplot2)
library(dplyr)
library(ggspatial)  # for north arrow & scale bar
library(ggnewscale)   
library(cowplot)
library(rnaturalearth)
library(rnaturalearthdata)
```


# Figure 1 - Map of BC fire centres and fire perimeters for 2023-2025
```{r}
fire_centres_path <- "data/raw/BC_Fire_Centres/DRP_MOF_FIRE_CENTRES_SP/DRPMFFRCNT_polygon.shp"
 
hist_fires_path   <- "data/raw/BC_Fire_Perimeters_Historical/PROT_HISTORICAL_FIRE_POLYS_SP/H_FIRE_PLY_polygon.shp"
 
curr_fires_path   <- "data/raw/BC_Fire_Perimeters_2025/PROT_CURRENT_FIRE_POLYS_SP/C_FIRE_PLY_polygon.shp"
 
```

Read in paths to shapefiles and load them as sf objects.
```{r}
fire_centres <- st_read(fire_centres_path, quiet = TRUE)
hist_fires   <- st_read(hist_fires_path, quiet = TRUE)
curr_fires   <- st_read(curr_fires_path, quiet = TRUE)
```

Filter historical fires to 2023 and 2024

```{r}
fires_2023 <- hist_fires |> 
  filter(FIRE_YEAR == 2023)

fires_2024 <- hist_fires |> 
  filter(FIRE_YEAR == 2024)

fires_2025 <- curr_fires
```

reproject to BC Albers (EPSG:3005)
```{r}
crs_bc <- 3005
 
fire_centres <- st_transform(fire_centres, crs_bc)
fires_2023   <- st_transform(fires_2023,   crs_bc)
fires_2024   <- st_transform(fires_2024,   crs_bc)
fires_2025   <- st_transform(fires_2025,   crs_bc)
```

colour palette


```{r}
centre_colours <- c(
  "Cariboo Fire Centre"      = "#F5E6C8",   # pale tan/cream
  "Coastal Fire Centre"      = "#F4C8A0",   # peach-orange
  "Kamloops Fire Centre"     = "#C8D8F0",   # light blue
  "Northwest Fire Centre"    = "#B8C8E8",   # muted blue
  "Prince George Fire Centre"= "#C8E0C8",   # light green
  "Southeast Fire Centre"    = "#D8C8E8"    # lavender
)
 
# Fire perimeter fills — bold, match original legend
fire_colours <- c(
  "2023" = "#FFD966",   # yellow
  "2024" = "#E08030",   # orange
  "2025" = "#C0281C"    # red
)
```

prepare fire layers with labels


```{r}
fires_2023$year <- "2023"
fires_2024$year <- "2024"
fires_2025$year <- "2025"
 
# Keep only geometry + year for a clean bind
fires_all <- bind_rows(
  fires_2023 %>% select(year) %>% st_as_sf(),
  fires_2024 %>% select(year) %>% st_as_sf(),
  fires_2025 %>% select(year) %>% st_as_sf()
)
 
fires_all$year <- factor(fires_all$year, levels = c("2023", "2024", "2025"))
```

Get fire centre name column


```{r}
names(fire_centres)
fire_centres <- fire_centres %>%
  rename(centre_name = MFFRCNTRNM)   
```

Build the map


```{r}
p_main <- ggplot() +
 
  geom_sf(
    data      = fire_centres,
    aes(fill  = centre_name),
    colour    = "gray60",
    linewidth = 0.4,
    alpha     = 0.85
  ) +
  scale_fill_manual(
    name   = "Fire Centres",
    values = centre_colours,
    guide  = guide_legend(
      order        = 2,
      override.aes = list(colour = "grey60", linewidth = 0.3)
    )
  ) +
 
  new_scale_fill() +
 
  geom_sf(
    data     = fires_all,
    aes(fill = year),
    colour   = "gray30",
    alpha    = 0.9
  ) +
  scale_fill_manual(
    name   = "Fires",
    values = fire_colours,
    labels = c("2023 perimeters", "2024 perimeters", "2025 perimeters"),
    guide  = guide_legend(
      order        = 1,
      override.aes = list(colour = NA)
    )
  ) +
 
  annotation_north_arrow(
    location    = "br",
    which_north = "true",
    pad_x       = unit(0.19, "in"),
    pad_y       = unit(0.2, "in"),
    style       = north_arrow_fancy_orienteering(
      fill      = c("grey30", "white"),
      line_col  = "grey30",
      text_col  = "grey30",
      text_size = 8
    )
  ) +
  annotation_scale(
    location   = "br",
    width_hint = 0.3,
    pad_x      = unit(0.5, "in"),
    pad_y      = unit(0.12, "in"),
    text_cex   = 0.7,
    line_col   = "grey30",
    text_col   = "grey30"
  ) +
 
  coord_sf(expand = FALSE) +

 
  theme_void(base_size = 11) +
  theme(
    plot.title           = element_text(face = "bold", size = 13,
                                        margin = margin(b = 6), hjust = 0),
    plot.caption         = element_text(colour = "grey50", size = 8,
                                        hjust = 1, margin = margin(t = 6)),
    plot.margin          = margin(12, 12, 12, 12),
    plot.background      = element_rect(fill = "transparent", colour = NA),
    panel.background     = element_rect(fill = "transparent", colour = NA),
    legend.position      = c(0.95, 0.97),
    legend.justification = c("right", "top"),
    legend.title         = element_text(face = "bold", size = 9),
    legend.text          = element_text(size = 9),
    legend.spacing.y     = unit(0.15, "cm"),
    legend.key.size      = unit(0.45, "cm"),
    legend.box           = "vertical",
    legend.box.spacing   = unit(0.1, "cm")
  )
 p_main
```


Inset map of Canada with BC highlighted

```{r}
canada    <- ne_states(country = "Canada", returnclass = "sf")
 
# Reproject to Canada Lambert Conformal Conic — looks good for the whole country
crs_canada <- "+proj=lcc +lat_1=49 +lat_2=77 +lat_0=63 +lon_0=-92 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
 
canada_proj  <- st_transform(canada, crs_canada)
 
# Highlight BC
bc_highlight <- canada_proj %>% filter(name == "British Columbia")
 
p_inset <- ggplot() +
 
  # Canadian provinces — all muted grey
  geom_sf(data = canada_proj, fill = "#D4D4D4", colour = "white", linewidth = 0.2) +
 
  # BC highlighted in a warm amber to stand out
  geom_sf(data = bc_highlight, fill = "#E08030", colour = "white", linewidth = 0.4) +
 
  # Crop to Canada + a bit of northern USA
  coord_sf(
    xlim   = c(-3e6, 3.2e6),
    ylim   = c(-1.5e6, 3.5e6),
    expand = FALSE,
    crs    = crs_canada
  ) +
 
  theme_void() +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    plot.margin      = margin(2, 2, 2, 2)
  )

 
```


```{r}
final_map <- ggdraw(p_main) +
  draw_plot(
    p_inset,
    x      = 0.01,   # left edge of inset (fraction of full plot width)
    y      = 0.1,   # bottom edge
    width  = 0.28,
    height = 0.22
  )
final_map

```



```{r}
ggsave(
  filename = "visualizations/maps/bc_fire_map.png",
  plot     = final_map,
  width    = 8,
  height   = 9,
  dpi      = 600,
  bg       = "transparent"
)
 

```

# Fire Map 2


```{r}
hist_fires_path  <- "data/raw/BC_Fire_Perimeters_Historical/PROT_HISTORICAL_FIRE_POLYS_SP/H_FIRE_PLY_polygon.shp"
curr_fires_path  <- "data/raw/BC_Fire_Perimeters_2025/PROT_CURRENT_FIRE_POLYS_SP/C_FIRE_PLY_polygon.shp"
study_zones_path <- "data/study_aoi/study_firezones.shp"
 
```


```{r}
hist_fires  <- st_read(hist_fires_path,  quiet = TRUE)
curr_fires  <- st_read(curr_fires_path,  quiet = TRUE)
study_zones <- st_read(study_zones_path, quiet = TRUE)
```



clip to zones
```{r}
study_union <- st_union(study_zones)   # single polygon for clipping
 
# 1950-2022: keep only fires fully WITHIN the study zones
clip_within <- function(x) {
  x <- st_make_valid(x)
  x[st_within(x, study_union, sparse = FALSE)[, 1], ]
}
 
# 2023-2025: keep fires that INTERSECT the study zones (then clip geometry)
# 2023-2025: keep fires that INTERSECT the study zones, but retain full geometry
clip_intersect <- function(x) {
  x <- st_make_valid(x)
  x[st_intersects(x, study_union, sparse = FALSE)[, 1], ]
}
 
 
fires_hist <- hist_fires %>% filter(FIRE_YEAR <= 2022) %>% clip_within()
fires_2023 <- hist_fires %>% filter(FIRE_YEAR == 2023) %>% clip_intersect()
fires_2024 <- hist_fires %>% filter(FIRE_YEAR == 2024) %>% clip_intersect()
fires_2025 <- curr_fires                               %>% clip_intersect()
```


```{r}
crs_bc <- 3005
 
study_zones <- st_transform(study_zones, crs_bc)
fires_hist  <- st_transform(fires_hist,  crs_bc)
fires_2023  <- st_transform(fires_2023,  crs_bc)
fires_2024  <- st_transform(fires_2024,  crs_bc)
fires_2025  <- st_transform(fires_2025,  crs_bc)
```


```{r}
study_bbox <- st_bbox(study_zones)
pad <- 60000   # metres padding around edge
xlim_bc <- c(study_bbox["xmin"] - pad, study_bbox["xmax"] + pad)
ylim_bc <- c(study_bbox["ymin"] - pad, study_bbox["ymax"] + pad)
 
```


```{r}
fire_colours <- c(
  "1950-2022" = "#9E9E9E",
  "2023"      = "#FFD966",
  "2024"      = "#E08030",
  "2025"      = "#C0281C"
)
 
# Two study zones — colours match the pink/lavender in your target map
# These will map to whatever the two MFFRZNNM values are
zone_colours <- c(
  "Fort St. John Fire Zone"    = "#E8D0E8",   # soft lavender-pink
  "Fort Nelson Fire Zone"  = "#C8E0C8"    # 
)
```


```{r}
fires_hist$period <- "1950-2022"
fires_2023$period <- "2023"
fires_2024$period <- "2024"
fires_2025$period <- "2025"
 
fires_all <- bind_rows(
  fires_hist %>% select(period) %>% st_as_sf(),
  fires_2023 %>% select(period) %>% st_as_sf(),
  fires_2024 %>% select(period) %>% st_as_sf(),
  fires_2025 %>% select(period) %>% st_as_sf()
) %>%
  mutate(period = factor(period, levels = c("1950-2022", "2023", "2024", "2025")))
```


```{r}
figure_2 <- ggplot() +
 
  # Study zones — distinct colours, clearly visible
  geom_sf(
    data      = study_zones,
    aes(fill  = MFFRZNNM),
    colour    = "grey60",
    linewidth = 0.5,
    alpha     = 0.6
  ) +
  scale_fill_manual(
    name   = "Fire Zones",
    values = zone_colours,
    guide  = guide_legend(
      order        = 2,
      override.aes = list(colour = "grey60", linewidth = 0.3, alpha = 0.8)
    )
  ) +
 
  new_scale_fill() +
 
  # Fire perimeters — historical grey first, recent on top
  geom_sf(
    data     = fires_all,
    aes(fill = period),
    colour   = NA,
    alpha    = 0.9
  ) +
  scale_fill_manual(
    name   = "Fires",
    values = fire_colours,
    labels = c("1950–2022 perimeters", "2023 perimeters",
               "2024 perimeters",      "2025 perimeters"),
    guide  = guide_legend(
      order        = 1,
      override.aes = list(colour = NA)
    )
  ) +
 
  # Zone name labels
  geom_sf_label(
    data          = study_zones,
    aes(label     = MFFRZNNM),
    size          = 3,
    colour        = "grey30",
    fill          = alpha("white", 0.6),
    label.size    = 0,
    label.padding = unit(0.15, "lines")
  ) +
 
  annotation_north_arrow(
    location    = "br",
    which_north = "true",
    pad_x       = unit(4, "in"),
    pad_y       = unit(1.2, "in"),
    style       = north_arrow_fancy_orienteering(
      fill      = c("grey30", "white"),
      line_col  = "grey30",
      text_col  = "grey30",
      text_size = 8
    )
  ) +
  annotation_scale(
    location   = "br",
    width_hint = 0.25,
    pad_x      = unit(4, "in"),
    pad_y      = unit(0.95, "in"),
    text_cex   = 0.7,
    line_col   = "grey30",
    text_col   = "grey30"
  ) +
 
  coord_sf(
    xlim   = xlim_bc,
    ylim   = ylim_bc,
    expand = FALSE
  )  +
 
  theme_void(base_size = 11) +
  theme(
    plot.title           = element_text(face = "bold", size = 13,
                                        margin = margin(b = 6), hjust = 0),
    plot.caption         = element_text(colour = "grey50", size = 8,
                                        hjust = 1, margin = margin(t = 6)),
    plot.margin          = margin(12, 12, 12, 12),
    plot.background      = element_rect(fill = "transparent", colour = NA),
    panel.background     = element_rect(fill = "transparent", colour = NA),
    legend.position      = c(0.25, 0.45),
    legend.justification = c("left", "top"),
    legend.background    = element_rect(fill = "white", colour = NA, linewidth = 0),
    legend.title         = element_text(face = "bold", size = 9),
    legend.text          = element_text(size = 8),
    legend.spacing.y     = unit(0.15, "cm"),
    legend.key.size      = unit(0.45, "cm"),
    legend.box           = "vertical",
    legend.box.spacing   = unit(0.1, "cm")
  )
figure_2
```


```{r}
ggsave(
  filename = "visualizations/maps/figure_2_fire_zones.png",
  plot     = figure_2,
  width    = 8,
  height   = 9,
  dpi      = 600,
  bg       = "transparent"
)
 
```

# Figure 3 - map of overwintering sites 2 and 3 year with 2023-2025 fire perimeters


```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggspatial)
library(stringr)
library(prettymapr)
```

```{r}
chains <- read_csv('results/multiyear_overwinter.csv')

perims_2023 <- st_read('data/processed_fire_perimeters/perims_2023_clipped_3005.shp', quiet = TRUE) |>st_make_valid() |> st_transform(4326)
perims_2024 <- st_read('data/processed_fire_perimeters/perims_2024_clipped_3005.shp', quiet = TRUE) |> st_make_valid() |> st_transform(4326)
perims_2025 <- st_read('data/processed_fire_perimeters/perims_2025_clipped_3005.shp', quiet = TRUE) |> st_make_valid() |> st_transform(4326)

# study zones
study_zones <- st_read('data/study_aoi/study_firezones.shp', quiet = TRUE) |> st_transform(4326)

# ignitions
ignitions_2024 <- st_read('results/ignitions_2024.geojson', quiet = TRUE) |> st_transform(4326)
ignitions_2025 <- st_read('results/ignitions_2025.geojson', quiet = TRUE) |> st_transform(4326)

clean_names <- function(df, year) {
  names(df) <- tolower(names(df))
  df <- df |>
    mutate(
      fire_id  = paste0(fire_year, '_', fire_no),
      geo_cat  = paste0('perim_', substr(as.character(year), 3, 4))
    )
  if (!'fire_cause' %in% names(df)) {
    df <- df |> mutate(fire_cause = NA_character_)
  }
  df |> select(fire_id, fire_year, fire_cause, size_ha, geo_cat, area_sqm, geometry)
}

perims_2023 <- clean_names(perims_2023, 2023)
perims_2024 <- clean_names(perims_2024, 2024)
perims_2025 <- clean_names(perims_2025, 2025)
```


```{r}
PERIM_PAL <- c(
  "2023 origin"        = "#FFD966",
  "2024 reactivation"  = "#E08030",
  "2025 reactivation"  = "#C0281C"
)
 
PERIM_ALPHA <- c(
  "Background"         = 0.30,
  "2023 origin"        = 0.75,
  "2024 reactivation"  = 0.75,
  "2025 reactivation"  = 0.75
)
 
HOTSPOT_PAL <- c(
  "2023\u21922024" = "#3079e0",
  "2024\u21922025" = "#930cb8"
)
 
theme_pub <- function() {
  theme_bw(base_size = 9, base_family = "sans") +
    theme(
      panel.grid        = element_blank(),
      panel.border      = element_rect(colour = NA, fill = NA),
      panel.background  = element_rect(fill = NA, colour = NA), 
      plot.background = element_rect(fill=NA, colour = NA), # Fort Nelson green
      axis.text.x  = element_blank(),
      axis.text.y  = element_blank(),
      axis.ticks   = element_blank(),
      axis.title = element_blank(),
      legend.title      = element_text(size = 7.5, face = "bold"),
      legend.text       = element_text(size = 7),
      legend.key        = element_rect(fill = NA),
      legend.key.size   = unit(0.4, "cm"),
      legend.background = element_rect(fill = "white", colour = "grey70",
                                       linewidth = 0.3),
      legend.margin     = margin(3, 5, 3, 5)
    )
}

zone_colours <- c(
  "Fort St. John Fire Zone"    = "#E8D0E8",   # soft lavender-pink
  "Fort Nelson Fire Zone"  = "#C8E0C8"    # 
)
```


```{r}
fires_crop <- bind_rows(
  perims_2023 |> filter(fire_id %in% chains$fire_id_2023) |>
    mutate(chain_class = '2023 origin'),
  perims_2024 |> filter(fire_id %in% chains$fire_id_2024) |>
    mutate(chain_class = '2024 reactivation'),
  perims_2025 |> filter(fire_id %in% chains$fire_id_2025) |>
    mutate(chain_class = '2025 reactivation')
)
```


```{r}
spring_sf <- bind_rows(
  ignitions_2024 |> filter(fire_id %in% chains$fire_id_2024) |>
    mutate(transition = '2023\u21922024'),
  ignitions_2025 |> filter(fire_id %in% chains$fire_id_2025) |>
    mutate(transition = '2024\u21922025')
)
```



```{r}
# ── compute per-chain bboxes ───────────────────────────────────────────────────
PAD <- 0.1

all_bboxes <- lapply(seq_len(nrow(chains)), function(i) {
  fires_crop |>
    filter(fire_id %in% c(chains$fire_id_2023[i],
                          chains$fire_id_2024[i],
                          chains$fire_id_2025[i])) |>
    st_bbox()
})

# maximum width and height across all chains
max_w <- max(sapply(all_bboxes, function(b) unname(b['xmax'] - b['xmin']))) + PAD * 2
max_h <- max(sapply(all_bboxes, function(b) unname(b['ymax'] - b['ymin']))) + PAD * 2

plot_list <- vector('list', nrow(chains))
bbox_list <- vector('list', nrow(chains))


for (i in seq_len(nrow(chains))) {

  row    <- chains[i, ]
  fid_23 <- row$fire_id_2023
  fid_24 <- row$fire_id_2024
  fid_25 <- row$fire_id_2025

  fires_chain  <- fires_crop |> filter(fire_id %in% c(fid_23, fid_24, fid_25))
  spring_chain <- spring_sf  |> filter(fire_id %in% c(fid_24, fid_25))

  # centre of this chain's fires
  b        <- all_bboxes[[i]]
  centre_x <- unname((b['xmin'] + b['xmax']) / 2)
  centre_y <- unname((b['ymin'] + b['ymax']) / 2)

  # apply same width and height to every plot
  bbox_pad <- st_bbox(
    c(xmin = centre_x - max_w / 2,
      ymin = centre_y - max_h / 2,
      xmax = centre_x + max_w / 2,
      ymax = centre_y + max_h / 2),
    crs = st_crs(4326)
  )

  crop_poly   <- st_as_sfc(bbox_pad)
  zones_chain <- study_zones |>
    st_crop(crop_poly) |>
    group_by(MFFRZNNM) |>
    summarise()

  p <- ggplot() +
    geom_sf(
      data      = zones_chain,
      aes(fill  = MFFRZNNM),
      colour    = 'grey60',
      linewidth = 0.5,
      alpha     = 0.5
    ) +
    scale_fill_manual(
      name   = 'Fire Zones',
      values = zone_colours,
      guide  = guide_legend(order = 2)
    ) +

    ggnewscale::new_scale_fill() +

    geom_sf(
      data      = fires_chain,
      aes(fill  = chain_class),
      colour    = 'grey40',
      linewidth = 0.3,
      alpha     = 0.6
    ) +
    scale_fill_manual(
      name   = 'Fire perimeter',
      values = PERIM_PAL,
      guide  = guide_legend(order = 1)
    ) +

    ggnewscale::new_scale_fill() +

    geom_sf(
      data  = spring_chain,
      aes(fill = transition),
      shape = 24, size = 2.5, colour = 'white', stroke = 0.4, alpha = 0.95
    ) +
    scale_fill_manual(
      name   = 'Overwintering Ignition',
      values = HOTSPOT_PAL,
      guide  = guide_legend(order = 3)
    ) +

    coord_sf(
      crs    = 4326,
      xlim   = c(bbox_pad['xmin'], bbox_pad['xmax']),
      ylim   = c(bbox_pad['ymin'], bbox_pad['ymax']),
      expand = FALSE
    ) +
    theme_pub() +
    theme(
  legend.position        = 'right',
  legend.justification   = c(0, 1),
  legend.box             = 'vertical',
  legend.box.background  = element_blank(),
  legend.box.margin      = margin(0, 0, 0, 0),
  legend.box.spacing     = unit(0.1, 'cm'),
  legend.background      = element_blank(),    # ← change this
  legend.key             = element_rect(fill = NA, colour = NA),
  plot.title             = element_text(size = 8, face = 'bold')
)
  bbox_list[[i]] <- bbox_pad

  plot_list[[i]] <- p
}

plot_list[[4]] + theme(legend.position = 'right')

```

p_1
```{r}
# pull chain 1 data
fid_23 <- chains$fire_id_2023[1]
fid_24 <- chains$fire_id_2024[1]
fid_25 <- chains$fire_id_2025[1]

fires_chain  <- fires_crop |> filter(fire_id %in% c(fid_23, fid_24, fid_25))
spring_chain <- spring_sf  |> filter(fire_id %in% c(fid_24, fid_25))

# tighter bbox — zoom to just the fires + small pad
bbox_zoom <- st_bbox(fires_chain)
bbox_zoom <- st_bbox(
  c(xmin = unname(bbox_zoom['xmin']) - 0.0005,
    ymin = unname(bbox_zoom['ymin']) - 0.0005,
    xmax = unname(bbox_zoom['xmax']) + 0.0005,
    ymax = unname(bbox_zoom['ymax']) + 0.0005),
  crs = st_crs(4326)
)

zones_zoom <- study_zones |>
  st_crop(st_as_sfc(bbox_zoom)) |>
  group_by(MFFRZNNM) |>
  summarise()

# replace plot 1 with zoomed version
plot_list[[1]] <- ggplot() +
  geom_sf(data = zones_zoom, aes(fill = MFFRZNNM),
          colour = 'grey60', linewidth = 0.5, alpha = 0.5) +
  scale_fill_manual(name = 'Fire Zones', values = zone_colours,
                    guide = guide_legend(order = 2)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = fires_chain, aes(fill = chain_class),
          colour = 'grey40', linewidth = 0.3, alpha = 0.6) +
  scale_fill_manual(name = 'Fire perimeter', values = PERIM_PAL,
                    guide = guide_legend(order = 1)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = spring_chain, aes(fill = transition),
          shape = 24, size = 2.5, colour = 'white', stroke = 0.4, alpha = 0.95) +
  scale_fill_manual(name = 'Overwintering Ignition', values = HOTSPOT_PAL,
                    guide = guide_legend(order = 3)) +
  coord_sf(crs = 4326,
           xlim = c(bbox_zoom['xmin'], bbox_zoom['xmax']),
           ylim = c(bbox_zoom['ymin'], bbox_zoom['ymax']),
           expand = FALSE) +
  theme_pub() +theme(
  legend.position        = 'right',
  legend.justification   = c(0, 1),
  legend.box             = 'vertical',
  legend.box.background  = element_blank(),
  legend.box.margin      = margin(0, 0, 0, 0),
  legend.box.spacing     = unit(0.1, 'cm'),
  legend.background      = element_blank(),    # ← change this
  legend.key             = element_rect(fill = NA, colour = NA),
  plot.title             = element_text(size = 8, face = 'bold')
)
p_1 <- plot_list[[1]] + theme(legend.position = 'right')

```

P_2

```{r}
# pull chain 2 data
fid_23 <- chains$fire_id_2023[2]
fid_24 <- chains$fire_id_2024[2]
fid_25 <- chains$fire_id_2025[2]

fires_chain  <- fires_crop |> filter(fire_id %in% c(fid_23, fid_24, fid_25))
spring_chain <- spring_sf  |> filter(fire_id %in% c(fid_24, fid_25))

# tighter bbox — zoom to just the fires + small pad
bbox_zoom <- st_bbox(fires_chain)
bbox_zoom <- st_bbox(
  c(xmin = unname(bbox_zoom['xmin']) - 0.0005,
    ymin = unname(bbox_zoom['ymin']) - 0.0005,
    xmax = unname(bbox_zoom['xmax']) + 0.0005,
    ymax = unname(bbox_zoom['ymax']) + 0.0005),
  crs = st_crs(4326)
)

zones_zoom <- study_zones |>
  st_crop(st_as_sfc(bbox_zoom)) |>
  group_by(MFFRZNNM) |>
  summarise()

# replace plot 2 with zoomed version
plot_list[[2]] <- ggplot() +
  geom_sf(data = zones_zoom, aes(fill = MFFRZNNM),
          colour = 'grey60', linewidth = 0.5, alpha = 0.5) +
  scale_fill_manual(name = 'Fire Zones', values = zone_colours,
                    guide = guide_legend(order = 2)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = fires_chain, aes(fill = chain_class),
          colour = 'grey40', linewidth = 0.3, alpha = 0.6) +
  scale_fill_manual(name = 'Fire perimeter', values = PERIM_PAL,
                    guide = guide_legend(order = 1)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = spring_chain, aes(fill = transition),
          shape = 24, size = 2.5, colour = 'white', stroke = 0.4, alpha = 0.95) +
  scale_fill_manual(name = 'Overwintering Ignition', values = HOTSPOT_PAL,
                    guide = guide_legend(order = 3)) +
  coord_sf(crs = 4326,
           xlim = c(bbox_zoom['xmin'], bbox_zoom['xmax']),
           ylim = c(bbox_zoom['ymin'], bbox_zoom['ymax']),
           expand = FALSE) +
  theme_pub() +theme(
  legend.position        = 'right',
  legend.justification   = c(0, 1),
  legend.box             = 'vertical',
  legend.box.background  = element_blank(),
  legend.box.margin      = margin(0, 0, 0, 0),
  legend.box.spacing     = unit(0.1, 'cm'),
  legend.background      = element_blank(),    # ← change this
  legend.key             = element_rect(fill = NA, colour = NA),
  plot.title             = element_text(size = 8, face = 'bold')
)
p_2 <- plot_list[[2]] + theme(legend.position = 'right')

```

P_3

```{r}
# pull chain 3 data
fid_23 <- chains$fire_id_2023[3]
fid_24 <- chains$fire_id_2024[3]
fid_25 <- chains$fire_id_2025[3]

fires_chain  <- fires_crop |> filter(fire_id %in% c(fid_23, fid_24, fid_25))
spring_chain <- spring_sf  |> filter(fire_id %in% c(fid_24, fid_25))

# tighter bbox — zoom to just the fires + small pad
bbox_zoom <- st_bbox(fires_chain)
bbox_zoom <- st_bbox(
  c(xmin = unname(bbox_zoom['xmin']) - 0.0005,
    ymin = unname(bbox_zoom['ymin']) - 0.0005,
    xmax = unname(bbox_zoom['xmax']) + 0.0005,
    ymax = unname(bbox_zoom['ymax']) + 0.0005),
  crs = st_crs(4326)
)

zones_zoom <- study_zones |>
  st_crop(st_as_sfc(bbox_zoom)) |>
  group_by(MFFRZNNM) |>
  summarise()

# replace plot 3 with zoomed version
plot_list[[3]] <- ggplot() +
  geom_sf(data = zones_zoom, aes(fill = MFFRZNNM),
          colour = 'grey60', linewidth = 0.5, alpha = 0.5) +
  scale_fill_manual(name = 'Fire Zones', values = zone_colours,
                    guide = guide_legend(order = 2)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = fires_chain, aes(fill = chain_class),
          colour = 'grey40', linewidth = 0.3, alpha = 0.6) +
  scale_fill_manual(name = 'Fire perimeter', values = PERIM_PAL,
                    guide = guide_legend(order = 1)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = spring_chain, aes(fill = transition),
          shape = 24, size = 2.5, colour = 'white', stroke = 0.4, alpha = 0.95) +
  scale_fill_manual(name = 'Overwintering Ignition', values = HOTSPOT_PAL,
                    guide = guide_legend(order = 3)) +
  coord_sf(crs = 4326,
           xlim = c(bbox_zoom['xmin'], bbox_zoom['xmax']),
           ylim = c(bbox_zoom['ymin'], bbox_zoom['ymax']),
           expand = FALSE) +
  theme_pub() +theme(
  legend.position        = 'right',
  legend.justification   = c(0, 1),
  legend.box             = 'vertical',
  legend.box.background  = element_blank(),
  legend.box.margin      = margin(0, 0, 0, 0),
  legend.box.spacing     = unit(0.1, 'cm'),
  legend.background      = element_blank(),    # ← change this
  legend.key             = element_rect(fill = NA, colour = NA),
  plot.title             = element_text(size = 8, face = 'bold')
)
p_3 <- plot_list[[3]] + theme(legend.position = 'right')

```

P_4
```{r}
# pull chain 4 data
fid_23 <- chains$fire_id_2023[4]
fid_24 <- chains$fire_id_2024[4]
fid_25 <- chains$fire_id_2025[4]

fires_chain  <- fires_crop |> filter(fire_id %in% c(fid_23, fid_24, fid_25))
spring_chain <- spring_sf  |> filter(fire_id %in% c(fid_24, fid_25))

# tighter bbox — zoom to just the fires + small pad
bbox_zoom <- st_bbox(fires_chain)
bbox_zoom <- st_bbox(
  c(xmin = unname(bbox_zoom['xmin']) - 0.0005,
    ymin = unname(bbox_zoom['ymin']) - 0.0005,
    xmax = unname(bbox_zoom['xmax']) + 0.0005,
    ymax = unname(bbox_zoom['ymax']) + 0.0005),
  crs = st_crs(4326)
)

zones_zoom <- study_zones |>
  st_crop(st_as_sfc(bbox_zoom)) |>
  group_by(MFFRZNNM) |>
  summarise()

# replace plot 4 with zoomed version
plot_list[[4]] <- ggplot() +
  geom_sf(data = zones_zoom, aes(fill = MFFRZNNM),
          colour = 'grey60', linewidth = 0.5, alpha = 0.5) +
  scale_fill_manual(name = 'Fire Zones', values = zone_colours,
                    guide = guide_legend(order = 2)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = fires_chain, aes(fill = chain_class),
          colour = 'grey40', linewidth = 0.3, alpha = 0.6) +
  scale_fill_manual(name = 'Fire perimeter', values = PERIM_PAL,
                    guide = guide_legend(order = 1)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = spring_chain, aes(fill = transition),
          shape = 24, size = 2.5, colour = 'white', stroke = 0.4, alpha = 0.95) +
  scale_fill_manual(name = 'Overwintering Ignition', values = HOTSPOT_PAL,
                    guide = guide_legend(order = 3)) +
  coord_sf(crs = 4326,
           xlim = c(bbox_zoom['xmin'], bbox_zoom['xmax']),
           ylim = c(bbox_zoom['ymin'], bbox_zoom['ymax']),
           expand = FALSE) +
  theme_pub() +theme(
  legend.position        = 'right',
  legend.justification   = c(0, 1),
  legend.box             = 'vertical',
  legend.box.background  = element_blank(),
  legend.box.margin      = margin(0, 0, 0, 0),
  legend.box.spacing     = unit(0.1, 'cm'),
  legend.background      = element_blank(),    # ← change this
  legend.key             = element_rect(fill = NA, colour = NA),
  plot.title             = element_text(size = 8, face = 'bold')
)
p_4 <- plot_list[[4]] + theme(legend.position = 'right')

```



```{r}
# pull chain 4 data
fid_23 <- chains$fire_id_2023[5]
fid_24 <- chains$fire_id_2024[5]
fid_25 <- chains$fire_id_2025[5]

fires_chain  <- fires_crop |> filter(fire_id %in% c(fid_23, fid_24, fid_25))
spring_chain <- spring_sf  |> filter(fire_id %in% c(fid_24, fid_25))

# tighter bbox — zoom to just the fires + small pad
bbox_zoom <- st_bbox(fires_chain)
bbox_zoom <- st_bbox(
  c(xmin = unname(bbox_zoom['xmin']) - 0.0005,
    ymin = unname(bbox_zoom['ymin']) - 0.0005,
    xmax = unname(bbox_zoom['xmax']) + 0.0005,
    ymax = unname(bbox_zoom['ymax']) + 0.0005),
  crs = st_crs(4326)
)

zones_zoom <- study_zones |>
  st_crop(st_as_sfc(bbox_zoom)) |>
  group_by(MFFRZNNM) |>
  summarise()

# replace plot 5 with zoomed version
plot_list[[5]] <- ggplot() +
  geom_sf(data = zones_zoom, aes(fill = MFFRZNNM),
          colour = 'grey60', linewidth = 0.5, alpha = 0.5) +
  scale_fill_manual(name = 'Fire Zones', values = zone_colours,
                    guide = guide_legend(order = 2)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = fires_chain, aes(fill = chain_class),
          colour = 'grey40', linewidth = 0.3, alpha = 0.6) +
  scale_fill_manual(name = 'Fire perimeter', values = PERIM_PAL,
                    guide = guide_legend(order = 1)) +
  ggnewscale::new_scale_fill() +
  geom_sf(data = spring_chain, aes(fill = transition),
          shape = 24, size = 2.5, colour = 'white', stroke = 0.4, alpha = 0.95) +
  scale_fill_manual(name = 'Overwintering Ignition', values = HOTSPOT_PAL,
                    guide = guide_legend(order = 3)) +
  coord_sf(crs = 4326,
           xlim = c(bbox_zoom['xmin'], bbox_zoom['xmax']),
           ylim = c(bbox_zoom['ymin'], bbox_zoom['ymax']),
           expand = FALSE) +
  theme_pub() +theme(
  legend.position        = 'right',
  legend.justification   = c(0, 1),
  legend.box             = 'vertical',
  legend.box.background  = element_blank(),
  legend.box.margin      = margin(0, 0, 0, 0),
  legend.box.spacing     = unit(0.1, 'cm'),
  legend.background      = element_blank(),    # ← change this
  legend.key             = element_rect(fill = NA, colour = NA),
  plot.title             = element_text(size = 8, face = 'bold')
)
p_5 <- plot_list[[5]] + theme(legend.position = 'right')
```



#### canvas


```{r}
library(cowplot)
library(ggplot2)
library(sf)
library(dplyr)

# ── Get centroid of each chain on the main map ─────────────────────────────────
# these are the arrow start points on the main map
chain_centroids <- lapply(seq_len(nrow(chains)), function(i) {
  fires_crop |>
    filter(fire_id %in% c(chains$fire_id_2023[i],
                          chains$fire_id_2024[i],
                          chains$fire_id_2025[i])) |>
    st_union() |>
    st_centroid() |>
    st_coordinates() |>
    as.data.frame() |>
    mutate(chain = i)
}) |> bind_rows()

# ── Build your main study area map (p_main = your Figure 2) ───────────────────
# add arrow endpoint dots on the main map
p_main_annotated <- figure_2 +
  geom_point(
    data  = chain_centroids,
    aes(x = X, y = Y),
    colour = 'black', size = 2, shape = 21, fill = 'white', stroke = 0.6
  ) +
  geom_text(
    data  = chain_centroids,
    aes(x = X, y = Y, label = letters[chain]),
    size  = 3, fontface = 'bold',
    nudge_x = 0.15, nudge_y = 0.1
  )

# ── Define inset positions (as fraction of canvas 0-1) ────────────────────────
# adjust these to get the layout you want
inset_w <- 0.22   # inset width as fraction of canvas
inset_h <- 0.28   # inset height as fraction of canvas

inset_positions <- list(
  list(x = 0.00, y = 0.65),   # a — top left
  list(x = 0.00, y = 0.35),   # b — mid left
  list(x = 0.00, y = 0.05),   # c — bottom left
  list(x = 0.78, y = 0.65),   # d — top right
  list(x = 0.78, y = 0.35)    # e — mid right
)

# ── Arrow endpoints: from inset edge to centroid on main map ───────────────────
# these define where arrows go FROM (inset side) and TO (map centroid)
# expressed as fractions of canvas width/height
# you will need to adjust these to match your map's coordinate space

# convert chain centroids from WGS84 to canvas fraction
# first get the main map's extent
main_xlim <- layer_scales(p_main)$x$get_limits()
main_ylim <- layer_scales(p_main)$y$get_limits()

to_canvas <- function(lon, lat, xlim, ylim) {
  list(
    x = (lon - xlim[1]) / diff(xlim),
    y = (lat - ylim[1]) / diff(ylim)
  )
}

# ── Assemble with cowplot ──────────────────────────────────────────────────────
plots_no_legend <- lapply(list(p_1, p_2, p_3, p_4, p_5), function(p) {
  p + theme(legend.position = 'none',
            plot.background = element_rect(fill = 'white', colour = 'grey40',
                                           linewidth = 0.4))
})

# start with main map as base
canvas <- ggdraw(p_main_annotated + theme(legend.position = 'none'))

# add each inset
for (i in seq_along(inset_positions)) {
  pos <- inset_positions[[i]]
  canvas <- canvas +
    draw_plot(
      plots_no_legend[[i]],
      x = pos$x, y = pos$y,
      width = inset_w, height = inset_h
    )
}

# add arrows connecting insets to chain locations on main map
# arrow FROM inset centre edge TO centroid on main map
arrow_data <- data.frame(
  x    = sapply(inset_positions, function(p) p$x + inset_w / 2),
  y    = sapply(inset_positions, function(p) p$y + inset_h / 2),
  xend = to_canvas(chain_centroids$X, chain_centroids$Y,
                   main_xlim, main_ylim)$x,
  yend = to_canvas(chain_centroids$X, chain_centroids$Y,
                   main_xlim, main_ylim)$y
)

canvas <- canvas +
  geom_segment(
    data = arrow_data,
    aes(x = x, y = y, xend = xend, yend = yend),
    arrow     = arrow(length = unit(0.2, 'cm'), type = 'closed'),
    colour    = 'grey20',
    linewidth = 0.5,
    inherit.aes = FALSE
  )

# add panel labels
for (i in seq_along(inset_positions)) {
  pos <- inset_positions[[i]]
  canvas <- canvas +
    draw_label(
      letters[i],
      x         = pos$x + 0.01,
      y         = pos$y + inset_h - 0.01,
      fontface  = 'bold',
      size      = 10,
      hjust     = 0,
      vjust     = 1
    )
}

canvas
```


```{r}
# ── transform centroids to EPSG:3005 ──────────────────────────────────────────
chain_centroids_3005 <- lapply(seq_len(nrow(chains)), function(i) {
  fires_crop |>st_transform(3005) |>   # use your EPSG:3005 version of fires_crop
    filter(fire_id %in% c(chains$fire_id_2023[i],
                          chains$fire_id_2024[i],
                          chains$fire_id_2025[i])) |>
    st_union() |>
    st_centroid() |>
    st_coordinates() |>
    as.data.frame() |>
    mutate(chain = i)
}) |> bind_rows()

# ── get main map axis limits from the built plot ───────────────────────────────
p_main_built <- ggplot_build(figure_2)
main_xlim    <- p_main_built$layout$panel_params[[1]]$x_range
main_ylim    <- p_main_built$layout$panel_params[[1]]$y_range

# ── convert EPSG:3005 coords to canvas fraction ───────────────────────────────
to_canvas_frac <- function(x, y, xlim, ylim) {
  data.frame(
    cx = (x - xlim[1]) / diff(xlim),
    cy = (y - ylim[1]) / diff(ylim)
  )
}

centroid_canvas <- to_canvas_frac(
  chain_centroids_3005$X,
  chain_centroids_3005$Y,
  main_xlim,
  main_ylim
)

# ── arrow FROM inset edge TO centroid ─────────────────────────────────────────
# inset edge: right edge for left insets, left edge for right insets
arrow_data <- data.frame(
  x    = c(
    inset_positions[[1]]$x + inset_w,       # a — right edge
    inset_positions[[2]]$x + inset_w,       # b — right edge
    inset_positions[[3]]$x + inset_w,       # c — right edge
    inset_positions[[4]]$x,                 # d — left edge
    inset_positions[[5]]$x                  # e — left edge
  ),
  y    = c(
    inset_positions[[1]]$y + inset_h / 2,
    inset_positions[[2]]$y + inset_h / 2,
    inset_positions[[3]]$y + inset_h / 2,
    inset_positions[[4]]$y + inset_h / 2,
    inset_positions[[5]]$y + inset_h / 2
  ),
  xend = centroid_canvas$cx,
  yend = centroid_canvas$cy
)

arrow_data[5, 'x']    <- inset_positions[[5]]$x + inset_w / 2   # bottom centre of inset
arrow_data[5, 'y']    <- inset_positions[[5]]$y                   # bottom edge
arrow_data[5, 'xend'] <- centroid_canvas$cx[5]
arrow_data[5, 'yend'] <- centroid_canvas$cy[5]
# ── rebuild canvas with corrected arrows ──────────────────────────────────────
canvas <- ggdraw(figure_2 + theme(legend.position = 'none'))

for (i in seq_along(inset_positions)) {
  pos    <- inset_positions[[i]]
  canvas <- canvas +
    draw_plot(
      plots_no_legend[[i]],
      x = pos$x, y = pos$y,
      width = inset_w, height = inset_h
    )
}

canvas <- canvas +
  geom_segment(
    data        = arrow_data,
    aes(x = x, y = y, xend = xend, yend = yend),
    arrow       = arrow(length = unit(0.2, 'cm'), type = 'closed'),
    colour      = 'grey20',
    linewidth   = 0.5,
    inherit.aes = FALSE
  )

for (i in seq_along(inset_positions)) {
  pos    <- inset_positions[[i]]
  canvas <- canvas +
    draw_label(
      letters[i],
      x        = pos$x + 0.01,
      y        = pos$y + inset_h - 0.01,
      fontface = 'bold', size = 10,
      hjust = 0, vjust = 1
    )
}

canvas
```


```{r}
library(svglite)

ggsave(
  filename = 'visualizations/maps/multiyear_chains.svg',
  plot     = canvas,
  width    = 12,
  height   = 8,
  units    = 'in',
  device   = svglite
)

```


```{r}
ggsave(
  filename = 'visualizations/maps/figure_2_fire_zones.svg',
  plot     = figure_2,
  width    = 12,
  height   = 8,
  units    = 'in',
  device   = svglite
)
```


```{r}
library(svglite)

svglite('visualizations/maps/legend_2.svg', width = 12, height = 8)
print(p_5)
dev.off()
```