

```{r}
library(sf)
library(ggplot2)
library(dplyr)
library(ggspatial)  # for north arrow & scale bar
library(ggnewscale)   
library(cowplot)
library(rnaturalearth)
library(rnaturalearthdata)
```


# Figure 1 - Map of BC fire centres and fire perimeters for 2023-2025
```{r}
fire_centres_path <- "data/raw/BC_Fire_Centres/DRP_MOF_FIRE_CENTRES_SP/DRPMFFRCNT_polygon.shp"
 
hist_fires_path   <- "data/raw/BC_Fire_Perimeters_Historical/PROT_HISTORICAL_FIRE_POLYS_SP/H_FIRE_PLY_polygon.shp"
 
curr_fires_path   <- "data/raw/BC_Fire_Perimeters_2025/PROT_CURRENT_FIRE_POLYS_SP/C_FIRE_PLY_polygon.shp"
 
```

Read in paths to shapefiles and load them as sf objects.
```{r}
fire_centres <- st_read(fire_centres_path, quiet = TRUE)
hist_fires   <- st_read(hist_fires_path, quiet = TRUE)
curr_fires   <- st_read(curr_fires_path, quiet = TRUE)
```

Filter historical fires to 2023 and 2024

```{r}
fires_2023 <- hist_fires |> 
  filter(FIRE_YEAR == 2023)

fires_2024 <- hist_fires |> 
  filter(FIRE_YEAR == 2024)

fires_2025 <- curr_fires
```

reproject to BC Albers (EPSG:3005)
```{r}
crs_bc <- 3005
 
fire_centres <- st_transform(fire_centres, crs_bc)
fires_2023   <- st_transform(fires_2023,   crs_bc)
fires_2024   <- st_transform(fires_2024,   crs_bc)
fires_2025   <- st_transform(fires_2025,   crs_bc)
```

colour palette


```{r}
centre_colours <- c(
  "Cariboo Fire Centre"      = "#F5E6C8",   # pale tan/cream
  "Coastal Fire Centre"      = "#F4C8A0",   # peach-orange
  "Kamloops Fire Centre"     = "#C8D8F0",   # light blue
  "Northwest Fire Centre"    = "#B8C8E8",   # muted blue
  "Prince George Fire Centre"= "#C8E0C8",   # light green
  "Southeast Fire Centre"    = "#D8C8E8"    # lavender
)
 
# Fire perimeter fills — bold, match original legend
fire_colours <- c(
  "2023" = "#FFD966",   # yellow
  "2024" = "#E08030",   # orange
  "2025" = "#C0281C"    # red
)
```

prepare fire layers with labels


```{r}
fires_2023$year <- "2023"
fires_2024$year <- "2024"
fires_2025$year <- "2025"
 
# Keep only geometry + year for a clean bind
fires_all <- bind_rows(
  fires_2023 %>% select(year) %>% st_as_sf(),
  fires_2024 %>% select(year) %>% st_as_sf(),
  fires_2025 %>% select(year) %>% st_as_sf()
)
 
fires_all$year <- factor(fires_all$year, levels = c("2023", "2024", "2025"))
```

Get fire centre name column


```{r}
names(fire_centres)
fire_centres <- fire_centres %>%
  rename(centre_name = MFFRCNTRNM)   
```

Build the map


```{r}
p_main <- ggplot() +
 
  geom_sf(
    data      = fire_centres,
    aes(fill  = centre_name),
    colour    = "gray60",
    linewidth = 0.4,
    alpha     = 0.85
  ) +
  scale_fill_manual(
    name   = "Fire Centres",
    values = centre_colours,
    guide  = guide_legend(
      order        = 2,
      override.aes = list(colour = "grey60", linewidth = 0.3)
    )
  ) +
 
  new_scale_fill() +
 
  geom_sf(
    data     = fires_all,
    aes(fill = year),
    colour   = "gray30",
    alpha    = 0.9
  ) +
  scale_fill_manual(
    name   = "Fires",
    values = fire_colours,
    labels = c("2023 perimeters", "2024 perimeters", "2025 perimeters"),
    guide  = guide_legend(
      order        = 1,
      override.aes = list(colour = NA)
    )
  ) +
 
  annotation_north_arrow(
    location    = "br",
    which_north = "true",
    pad_x       = unit(0.19, "in"),
    pad_y       = unit(0.2, "in"),
    style       = north_arrow_fancy_orienteering(
      fill      = c("grey30", "white"),
      line_col  = "grey30",
      text_col  = "grey30",
      text_size = 8
    )
  ) +
  annotation_scale(
    location   = "br",
    width_hint = 0.3,
    pad_x      = unit(0.5, "in"),
    pad_y      = unit(0.12, "in"),
    text_cex   = 0.7,
    line_col   = "grey30",
    text_col   = "grey30"
  ) +
 
  coord_sf(expand = FALSE) +

 
  theme_void(base_size = 11) +
  theme(
    plot.title           = element_text(face = "bold", size = 13,
                                        margin = margin(b = 6), hjust = 0),
    plot.caption         = element_text(colour = "grey50", size = 8,
                                        hjust = 1, margin = margin(t = 6)),
    plot.margin          = margin(12, 12, 12, 12),
    plot.background      = element_rect(fill = "transparent", colour = NA),
    panel.background     = element_rect(fill = "transparent", colour = NA),
    legend.position      = c(0.95, 0.97),
    legend.justification = c("right", "top"),
    legend.title         = element_text(face = "bold", size = 9),
    legend.text          = element_text(size = 9),
    legend.spacing.y     = unit(0.15, "cm"),
    legend.key.size      = unit(0.45, "cm"),
    legend.box           = "vertical",
    legend.box.spacing   = unit(0.1, "cm")
  )
 p_main
```


Inset map of Canada with BC highlighted

```{r}
canada    <- ne_states(country = "Canada", returnclass = "sf")
 
# Reproject to Canada Lambert Conformal Conic — looks good for the whole country
crs_canada <- "+proj=lcc +lat_1=49 +lat_2=77 +lat_0=63 +lon_0=-92 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs"
 
canada_proj  <- st_transform(canada, crs_canada)
 
# Highlight BC
bc_highlight <- canada_proj %>% filter(name == "British Columbia")
 
p_inset <- ggplot() +
 
  # Canadian provinces — all muted grey
  geom_sf(data = canada_proj, fill = "#D4D4D4", colour = "white", linewidth = 0.2) +
 
  # BC highlighted in a warm amber to stand out
  geom_sf(data = bc_highlight, fill = "#E08030", colour = "white", linewidth = 0.4) +
 
  # Crop to Canada + a bit of northern USA
  coord_sf(
    xlim   = c(-3e6, 3.2e6),
    ylim   = c(-1.5e6, 3.5e6),
    expand = FALSE,
    crs    = crs_canada
  ) +
 
  theme_void() +
  theme(
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    plot.margin      = margin(2, 2, 2, 2)
  )

 
```


```{r}
final_map <- ggdraw(p_main) +
  draw_plot(
    p_inset,
    x      = 0.01,   # left edge of inset (fraction of full plot width)
    y      = 0.1,   # bottom edge
    width  = 0.28,
    height = 0.22
  )
final_map

```



```{r}
ggsave(
  filename = "visualizations/maps/bc_fire_map.png",
  plot     = final_map,
  width    = 8,
  height   = 9,
  dpi      = 600,
  bg       = "transparent"
)
 

```

# Fire Map 2


```{r}
hist_fires_path  <- "data/raw/BC_Fire_Perimeters_Historical/PROT_HISTORICAL_FIRE_POLYS_SP/H_FIRE_PLY_polygon.shp"
curr_fires_path  <- "data/raw/BC_Fire_Perimeters_2025/PROT_CURRENT_FIRE_POLYS_SP/C_FIRE_PLY_polygon.shp"
study_zones_path <- "data/study_aoi/study_firezones.shp"
 
```


```{r}
hist_fires  <- st_read(hist_fires_path,  quiet = TRUE)
curr_fires  <- st_read(curr_fires_path,  quiet = TRUE)
study_zones <- st_read(study_zones_path, quiet = TRUE)
```



clip to zones
```{r}
study_union <- st_union(study_zones)   # single polygon for clipping
 
# 1950-2022: keep only fires fully WITHIN the study zones
clip_within <- function(x) {
  x <- st_make_valid(x)
  x[st_within(x, study_union, sparse = FALSE)[, 1], ]
}
 
# 2023-2025: keep fires that INTERSECT the study zones (then clip geometry)
# 2023-2025: keep fires that INTERSECT the study zones, but retain full geometry
clip_intersect <- function(x) {
  x <- st_make_valid(x)
  x[st_intersects(x, study_union, sparse = FALSE)[, 1], ]
}
 
 
fires_hist <- hist_fires %>% filter(FIRE_YEAR <= 2022) %>% clip_within()
fires_2023 <- hist_fires %>% filter(FIRE_YEAR == 2023) %>% clip_intersect()
fires_2024 <- hist_fires %>% filter(FIRE_YEAR == 2024) %>% clip_intersect()
fires_2025 <- curr_fires                               %>% clip_intersect()
```


```{r}
crs_bc <- 3005
 
study_zones <- st_transform(study_zones, crs_bc)
fires_hist  <- st_transform(fires_hist,  crs_bc)
fires_2023  <- st_transform(fires_2023,  crs_bc)
fires_2024  <- st_transform(fires_2024,  crs_bc)
fires_2025  <- st_transform(fires_2025,  crs_bc)
```


```{r}
study_bbox <- st_bbox(study_zones)
pad <- 60000   # metres padding around edge
xlim_bc <- c(study_bbox["xmin"] - pad, study_bbox["xmax"] + pad)
ylim_bc <- c(study_bbox["ymin"] - pad, study_bbox["ymax"] + pad)
 
```


```{r}
fire_colours <- c(
  "1950-2022" = "#9E9E9E",
  "2023"      = "#FFD966",
  "2024"      = "#E08030",
  "2025"      = "#C0281C"
)
 
# Two study zones — colours match the pink/lavender in your target map
# These will map to whatever the two MFFRZNNM values are
zone_colours <- c(
  "Fort St. John Fire Zone"    = "#E8D0E8",   # soft lavender-pink
  "Fort Nelson Fire Zone"  = "#C8E0C8"    # slightly warmer pink
)
```


```{r}
fires_hist$period <- "1950-2022"
fires_2023$period <- "2023"
fires_2024$period <- "2024"
fires_2025$period <- "2025"
 
fires_all <- bind_rows(
  fires_hist %>% select(period) %>% st_as_sf(),
  fires_2023 %>% select(period) %>% st_as_sf(),
  fires_2024 %>% select(period) %>% st_as_sf(),
  fires_2025 %>% select(period) %>% st_as_sf()
) %>%
  mutate(period = factor(period, levels = c("1950-2022", "2023", "2024", "2025")))
```


```{r}
figure_2 <- ggplot() +
 
  # Study zones — distinct colours, clearly visible
  geom_sf(
    data      = study_zones,
    aes(fill  = MFFRZNNM),
    colour    = "grey60",
    linewidth = 0.5,
    alpha     = 0.6
  ) +
  scale_fill_manual(
    name   = "Fire Zones",
    values = zone_colours,
    guide  = guide_legend(
      order        = 2,
      override.aes = list(colour = "grey60", linewidth = 0.3, alpha = 0.8)
    )
  ) +
 
  new_scale_fill() +
 
  # Fire perimeters — historical grey first, recent on top
  geom_sf(
    data     = fires_all,
    aes(fill = period),
    colour   = NA,
    alpha    = 0.9
  ) +
  scale_fill_manual(
    name   = "Fires",
    values = fire_colours,
    labels = c("1950–2022 perimeters", "2023 perimeters",
               "2024 perimeters",      "2025 perimeters"),
    guide  = guide_legend(
      order        = 1,
      override.aes = list(colour = NA)
    )
  ) +
 
  # Zone name labels
  geom_sf_label(
    data          = study_zones,
    aes(label     = MFFRZNNM),
    size          = 3,
    colour        = "grey30",
    fill          = alpha("white", 0.6),
    label.size    = 0,
    label.padding = unit(0.15, "lines")
  ) +
 
  annotation_north_arrow(
    location    = "br",
    which_north = "true",
    pad_x       = unit(4, "in"),
    pad_y       = unit(1.2, "in"),
    style       = north_arrow_fancy_orienteering(
      fill      = c("grey30", "white"),
      line_col  = "grey30",
      text_col  = "grey30",
      text_size = 8
    )
  ) +
  annotation_scale(
    location   = "br",
    width_hint = 0.25,
    pad_x      = unit(4, "in"),
    pad_y      = unit(0.95, "in"),
    text_cex   = 0.7,
    line_col   = "grey30",
    text_col   = "grey30"
  ) +
 
  coord_sf(
    xlim   = xlim_bc,
    ylim   = ylim_bc,
    expand = FALSE
  )  +
 
  theme_void(base_size = 11) +
  theme(
    plot.title           = element_text(face = "bold", size = 13,
                                        margin = margin(b = 6), hjust = 0),
    plot.caption         = element_text(colour = "grey50", size = 8,
                                        hjust = 1, margin = margin(t = 6)),
    plot.margin          = margin(12, 12, 12, 12),
    plot.background      = element_rect(fill = "transparent", colour = NA),
    panel.background     = element_rect(fill = "transparent", colour = NA),
    legend.position      = c(0.25, 0.45),
    legend.justification = c("left", "top"),
    legend.background    = element_rect(fill = "white", colour = NA, linewidth = 0),
    legend.title         = element_text(face = "bold", size = 9),
    legend.text          = element_text(size = 8),
    legend.spacing.y     = unit(0.15, "cm"),
    legend.key.size      = unit(0.45, "cm"),
    legend.box           = "vertical",
    legend.box.spacing   = unit(0.1, "cm")
  )
figure_2
```


```{r}
ggsave(
  filename = "visualizations/maps/figure_2_fire_zones.png",
  plot     = figure_2,
  width    = 8,
  height   = 9,
  dpi      = 600,
  bg       = "transparent"
)
 
```

# Figure 3 - map of overwintering sites 2 and 3 year with 2023-2025 fire perimeters
Overlap in 